### 2023-10-14 chapters/paper_automated.tex
Fix many spelling and language errors

\section*{Abstract}
Medical coding is the task of assigning medical codes to clinical free-text documentation.
Healthcare professionals manually assign such codes to track patient diagnoses and treatments. Automated medical coding can considerably alleviate this administrative burden. In this paper, we reproduce, compare, and analyze state-of-the-art automated medical coding machine learning models. We show that several models underperform due to weak configurations, poorly sampled train-test splits, and insufficient evaluation. In previous work, the macro F1-score has been calculated sub-optimally, and our correction doubles it. We contribute a revised model comparison using stratified sampling and identical experimental setups, including hyperparameters and decision boundary tuning. We analyze prediction errors to validate and falsify assumptions of previous works. The analysis confirms that all models struggle with rare codes, while long documents only have a negligible impact. Finally, we present the first comprehensive results on the newly released MIMIC-IV dataset using the reproduced models. We release our code, model parameters, and new MIMIC-III and MIMIC-IV training and evaluation pipelines to accommodate fair future comparisons.\footnote{\label{footnote:source_code}\url{https://github.com/JoakimEdin/medical-coding-reproducibility}}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{paper_automated/mimiciv_code_freq.png}
    \caption[The frequency of ICD-9 and ICD-10 codes in MIMIC-IV before pre-processing.]{The frequency of ICD-9 and ICD-10 codes in MIMIC-IV before pre-processing. As discussed in \cref{sec: splits}, we removed codes with fewer than ten instances (dashed line).}
    \label{fig:mimiciv_code_dist}
\end{figure}

@@

\textcite{mullenbachExplainablePredictionMedical2018} released code for pre-processing the discharge summaries, generating the train-test split, and evaluating model performance on MIMIC-III which many subsequent papers have used \parencite{vuLabelAttentionModel2020, huangPLMICDAutomaticICD2022, liICDCodingClinical2020, baoMedicalCodePrediction2021, yuanCodeSynonymsMatter2022, kimReadAttendCode2021}.
Pre-processing consisted of lowercasing all text and removing words that only contain out-of-alphabet characters. Predicting procedures and diagnosis codes were treated as a single task.
The dataset was split into training, validation, and test sets using random sampling, ensuring that no patient occurred in both the training and test set. The (non-stratified) random sampling lead to 54\% of the ICD codes in MIMIC-III \textit{full} not being sampled in the test set. This complicates the interpretation of results since these codes only contribute true negatives or false positives. 
Models are evaluated using the micro and macro average of the area under the curve of the receiver operating characteristics (AUC-ROC), F1-score, and precision@k.

While most papers use the pre-processing, train-test split, and evaluation metrics described above, they differ in several aspects of training. This may lead to performance differences unrelated to modeling choices which are undesirable when seeking to compare models. 
For instance, due to varying memory constraints of different models, documents are usually truncated to some maximum length. In the literature, this maximum varies from 2,500 to 4,000 words \parencite{mullenbachExplainablePredictionMedical2018, vuLabelAttentionModel2020, huangPLMICDAutomaticICD2022}.
Furthermore, not all papers tune the prediction decision boundary but simply set it to 0.5,
 hyperparameter search ranges and sampling methods  vary between works, and learning rate schedulers are only used in LAAT and PLM-ICD\parencite{mullenbachExplainablePredictionMedical2018, liICDCodingClinical2020}. In LAAT, the learning rate was decreased by 90\% when the F1-score had not increased for five epochs. PLM-ICD used a schedule with linear warmup followed by linear decay.

@@

\begin{sidewaystable}
    \centering
    \caption[Hyperparameters, maximum document lengths, and decision boundary tuning strategies used in the original works compared to improved settings.]{Hyperparameters, maximum document lengths, and decision boundary tuning strategies used in the original works compared to the optimal settings found in this paper (marked with *). LR is the learning rate scheduler. ``Length" is the maximum number of words a document can contain before being truncated. $\dagger$ applies to models using word-piece tokenization. These models were filtered on the number of sub-words instead of full words. ``DB tune" is whether the optimal decision boundary was found using the validation set. If a paper did not tune the decision boundary, it was set to 0.5.}
    \label{tab:hyperparameters}
    \resizebox*{\textwidth}{!}{%
    \begin{tabular}{lccccccccc}
        \toprule
        & \multicolumn{7}{c}{Hyperparameters} & & \\
        \cmidrule(lr){2-8}
        Model & Batch Size & Weight Decay & Learning Rate & Dropout & LR Scheduler & Optimizer & Epochs & Length & DB tune \\
        \midrule
        Bi-GRU  & 16 &  0.0 & 0.003 &  0.2   & no  & Adam &  100 & 2500 & no\\
        Bi-GRU$^*$  & 8 &  0.0001 & 0.001 &  0   & yes  & AdamW &  20 & 4000 & yes\\
        \hline
        CNN  & 16 &  0.0 & 0.003  &  0.2 & no  & Adam &  100 & 2500 & no\\
        CNN$^*$    & 8 &  0.00001 & 0.001  &  0 & yes  & AdamW &  20 & 4000 & yes \\
        \hline
        CAML  & 16 &  0.0 & 0.0001  &  0.2  & no  & Adam  &  200 & 2500 & no\\
        CAML$^*$    & 8 &  0.001 & 0.005  &  0.2  & yes  & AdamW  &  20 & 4000 & yes\\
        \hline
        MultiResCNN & 16 &  0.0 & 0.0001  &  0.2  & no  & Adam &  200 & 2500 & no\\
        MultiResCNN$^*$  & 16 &  0.0001 & 0.0005  &  0.2  & yes  & AdamW &  20 & 4000 & yes\\
        \hline
        LAAT   & 8 &  0.0 & 0.0001  &  0.3  & yes  & AdamW &  50 & 4000 & no\\
        LAAT$^*$    & 8 &  0.001 & 0.001  &  0.2  & yes  & AdamW &  20 & 4000 & yes\\
        \hline
        PLM-ICD  & 8 &  0.0 & 0.00005  &  0.2  & yes  & AdamW &  20 & 3072$^\dagger$ & yes \\
        PLM-ICD$^*$   & 16 &  0.0 & 0.00005  &  0.2  & yes  & AdamW &  20 & 4000 & yes \\
        \bottomrule
    \end{tabular}%
    }
\end{sidewaystable}

Previous works calculated the macro F1-score as the harmonic mean of the macro precision and macro recall \parencite{mullenbachExplainablePredictionMedical2018, liICDCodingClinical2020, vuLabelAttentionModel2020, huangPLMICDAutomaticICD2022}. \textcite{opitzMacroF1Macro2021} analyze macro F1 formulas common in multi-class and multi-label classification. They demonstrate that the above formulation is sub-optimal, as it rewards heavily biased classifiers in unbalanced datasets. Therefore, as recommended by the authors, we calculated the macro F1-score as the arithmetic mean of the F1-score for each class.
As seen in \cref{tab:subsets}, 54\% of codes in MIMIC-III \textit{full} are missing in the test set. Previous works set the F1-score of all the missing codes in the test set to 0, resulting in a misleadingly low macro F1-score. Because 54\% of the codes are missing, the maximum possible macro F1-score is 46\%. 
We ignored all codes not in the test set for our reproduction, essentially trading bias for variance. 
For our revised comparison, we resolved the issue by instead sampling new splits that reduce missing codes to a negligible fraction (see \cref{sec: splits}) and ignoring the few that were still missing.

@@

\subsection{Revised comparison}
To address the issues associated with comparing results reported by previous works described in \cref{sec: training and evaluation methods,sec:metrics}, we perform a revised model comparison. We run experiments on the new MIMIC-III \textit{clean}, MIMIC-IV \textit{ICD-9}, and \textit{ICD-10}.
All models were trained for 20 epochs using a learning rate schedule with linear warmup for the first 2K updates followed by linear decay \parencite{huangPLMICDAutomaticICD2022}. We found this schedule to speed up the training convergence of all the models.
Whereas original works use Adam or AdamW, we used AdamW for all experiments as it corrects the weight decay implementation of Adam \parencite{kingmaAdamMethodStochastic2017, loshchilovDecoupledWeightDecay2022}. For each model, we tuned the decision boundary to maximize the micro F1-score on the validation set. We used randomized sampling to find optimal settings for dropout, weight decay, learning rate, and batch size. The hyperparameter search was performed on MIMIC-III \textit{clean}, and the MIMIC-IV splits. We found that the best setting for each model generalized across datasets. Using this setting, we ran each model ten times with different seeds on each dataset. All documents were truncated to a maximum of 4000 words. The hyperparameters, maximum document lengths, and decision boundary tuning strategy are summarised in \cref{tab:hyperparameters}.

@@

\begin{sidewaystable}
    \centering
    \caption[Reproduced test set results compared with those from the original works.]{Reproduced test set results compared with those from the original works. Our reproduced results are indicated with *. The results were reproduced on MIMIC-III v1.4 with the preprocessing pipeline and splits of \textcite{mullenbachExplainablePredictionMedical2018}. Each model was reproduced using the hyperparameters presented in the respective paper. We use both macro F1 formulas: Macro$^\dagger$ refers to the method used in the original work, while Macro refers to the corrected version used in this paper.}
    \label{tab:reproduced_results}
    \resizebox*{\textwidth}{!}{%
    \begin{tabular}{l  cc  ccc  cc  cc  ccc  c}
        \toprule
        & \multicolumn{7}{c}{MIMIC-III \textit{full}} & \multicolumn{6}{c}{MIMIC-III \textit{50}} \\
        \cmidrule(lr){2-8}\cmidrule(lr){9-14}
        & \multicolumn{2}{c}{AUC-ROC} & \multicolumn{3}{c}{F1} & \multicolumn{2}{c}{Precision@$k$}  & \multicolumn{2}{c}{AUC-ROC} & \multicolumn{3}{c}{F1} & Precision@$k$  \\
        & Micro & Macro & Micro & Macro$^\dagger$ & Macro & 8 & 15 & Micro & Macro & Micro & Macro$^\dagger$ & Macro & 5 \\
        \midrule
        CNN &  96.9 & 80.6 &  41.9& 4.2 & -& 58.1 &  48.8  & 90.7 & 87.6 & 62.5 & 57.6 & - & 62.0  \\
        CNN$^*$ & 97.3 & 83.1 & 41.5 & 3.4 & 6.7 & 61.9 & 47.2 & 91.9 & 89.2 & 64.9 & 58.8 & 58.0 &  62.6 \\
        \hline
        Bi-GRU & 97.1 & 82.2 & 41.7 & 3.8 & - & 58.5 & 44.5  &  89.2 & 82.8 & 54.9 &  48.4 & - &  59.1  \\
        Bi-GRU$^*$  & 98.0 & 87.1 & 42.6 & 3.6 & 7.0 & 65.0 & 49.8 & 89.3 & 85.2 & 56.1 & 46.2 & 43.1 & 57.9  \\
        \hline
        CAML  &  98.6 &  89.5 &  53.9 & 8.8 & - &  70.9 &  56.1  & 90.9 & 87.5 &  61.4 &  53.2 & - &   60.9  \\
        CAML$^*$ & 98.4 & 88.4 & 49.5 & 5.6 & 11.3 & 69.9 & 54.9 & 91.1 & 87.5 & 60.6 & 52.4 & 51.0 & 61.1  \\
        \hline
        MultiResCNN  &  98.6 &  91.0 & 55.2 & 8.6 & - & 73.4 & 58.4  &  93.8 &  89.9 &  67.0 &  60.6 & - &  64.1 \\
        MultiResCNN$^*$  & 98.6 & 90.8 & 56.5 & 9.2 & 18.5 & 73.4 & 58.4 & 92.4 & 89.7 & 67.3 & 62.2 & 61.1 & 63.4  \\
        \hline
        LAAT  &  98.8 &  91.9 &  57.5 & 9.9 & - &  74.5 &  59.1  &  94.6 &  92.5 &  71.5 &  66.6 & - &  67.5 \\
        LAAT$^*$ & 98.6 & 89.5 & 56.1 & 8.2 & 16.2 & 73.9 & 58.7 & 92.8 & 90.5 & 66.8 & 60.8 & 59.2 & 64.0  \\
        \hline
        PLM-ICD  &  98.9 & 92.6 &  59.8 & 10.4 & - &  77.1 &  61.3  & - & - & - & - & - & -  \\
        PLM-ICD$^*$ & 98.8 & 92.3 & 58.9 & 11.1 & 22.8 & 75.7 & 60.5 & 93.8 & 91.7 & 70.5 & 66.3 & 65.4 & 65.7 \\
        \bottomrule
    \end{tabular}%
    }
\end{sidewaystable}

@@

    \caption[Results on the MIMIC-III \textit{clean}, MIMIC-IV \textit{ICD-9} and MIMIC-IV  \textit{ICD-10} test sets presented as percentages.]{Results on the MIMIC-III \textit{clean}, MIMIC-IV \textit{ICD-9} and MIMIC-IV  \textit{ICD-10} test sets presented as percentages. Micro F1-scores rank the table in ascending order. Each model was trained ten times with different seeds. We performed a McNemar's test with Bonferroni correction and found that all the models are significantly different ($p < 0.001$).}

@@

\begin{sidewaystable}
    \centering
    \caption[Ablation study on MIMIC-III \textit{clean}.]{Ablation study on MIMIC-III \textit{clean}. The numbers are the micro/macro F1-scores on the test set.}
    \label{tab:ablation_study}
    \resizebox*{\textwidth}{!}{%
    \begin{tabular}{lcccccc}
        \toprule
         & PLM-ICD & LAAT & MultiResCNN & CAML & Bi-GRU & CNN \\
        \midrule
        Our result & 59.6/26.6 & 57.8/22.6 & 56.4/22.9 & 55.4/20.4 & 49.7/12.2 & 48.0/9.9 \\
        \hline
        Input length truncated at 2500 words & 59.4/26.2 & 57.6/22.3 & 56.0/23.2 & 54.8/19.7 & 49.4/12.0 & 47.9/9.8 \\
        No decision boundary tuning & 58.7/23.0 & 56.2/19.0 & 56.2/22.6 & 53.3/17.1 & 45.3/8.1 & 43.8/7.0 \\
        Original hyperparameters & 59.6/27.0 & 57.5/21.6 & 56.4/20.0 & 52.8/17.3 & 48.1/11.2 & 46.9/10.2 \\
        \bottomrule
        \end{tabular}%
    }
\end{sidewaystable}

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.9\textwidth}
        \centering
        \includegraphics[width=\textwidth]{paper_automated/threshold_tuning_micro_full.png}
        % \caption{}
        % \label{fig:threshold_tuning_micro}
    \end{subfigure}
    \begin{subfigure}[b]{0.9\textwidth}
        \centering
        \includegraphics[width=\textwidth]{paper_automated/threshold_tuning_macro_full.png}
        % \caption{}
        % \label{fig:threshold_tuning_macro}
    \end{subfigure}%
    \caption[Relationship between chosen threshold and F1-score of every reproduced model.]{The relationship between chosen threshold and F1-score of every reproduced model in \cref{tab:reproduced_results}. The left figure shows the micro F1-score, and the right shows the macro F1-score. The models were evaluated on MIMIC-III \textit{clean}.}
    \label{fig:threshold_tuning}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.95\textwidth}
        \centering
        \includegraphics[width=\textwidth]{paper_automated/train_size_f1_micro.png}
        % \caption{}
        % \label{fig:train_size_micro}
    \end{subfigure}
    \begin{subfigure}[b]{0.95\textwidth}
        \centering
        \includegraphics[width=\textwidth]{paper_automated/train_size_f1_macro.png}
        % \caption{}
        % \label{fig:train_size_macro}
    \end{subfigure}
    \caption[The relationship between the number of training examples and F1-score on MIMIC-IV \textit{ICD-9}.]{The relationship between the number of training examples and F1-score on MIMIC-IV \textit{ICD-9}. The left figure shows the F1 Micro score on the y-axis, while the right figure shows the F1 Macro score.}
    \label{fig:train_size}
\end{figure}

@@

\begin{table}
    \centering
    \caption[Correlation between the F1-score and the logarithm of code frequency and document length on MIMIC-IV \textit{ICD-9}.]{Correlation between the F1-score and the logarithm of code frequency and document length on MIMIC-IV \textit{ICD-9}. As discussed in \cref{subsubsec:document-length}, we only considered document lengths between 1000 and 4000 words. All correlations are statistically significant ($p<0.001$).}
    \label{tab:correlations}
    \begin{tabular}{lcccc}
        \toprule
        & \multicolumn{2}{c}{Code frequency} & \multicolumn{2}{c}{Document lengths}\\
        \cmidrule(lr){2-3}\cmidrule(lr){4-5}
        & Pearson & Spearman & Pearson & Spearman \\
        \midrule
        CNN  & 0.61 & 0.68 & -0.09 & -0.08 \\
        Bi-GRU & 0.57 & 0.65 & -0.08 & -0.07 \\
        CAML & 0.56 & 0.60 & -0.03 & -0.03   \\
        MultiResCNN & 0.47 & 0.53 & -0.02 & -0.03  \\
        LAAT & 0.52 & 0.57 & -0.02 & -0.02 \\
        PLM-ICD & 0.48 & 0.52 & -0.02 & -0.02   \\
        \bottomrule
    \end{tabular}
\end{table}

>>>>

\section*{Abstract}
Medical coding is the task of assigning medical codes to clinical free-text documentation.
Healthcare professionals manually assign such codes to track patient diagnoses and treatments. Automated medical coding can considerably alleviate this administrative burden. In this paper, we reproduce, compare, and analyze state-of-the-art automated medical coding machine learning models. We show that several models underperform due to weak configurations, poorly sampled train-test splits, and insufficient evaluation. In previous work, the macro F1-score has been calculated suboptimally, and our correction doubles it. We contribute a revised model comparison using stratified sampling and identical experimental setups, including hyperparameters and decision boundary tuning. We analyze prediction errors to validate and falsify assumptions of previous works. The analysis confirms that all models struggle with rare codes, while long documents only have a negligible impact. Finally, we present the first comprehensive results on the newly released MIMIC-IV dataset using the reproduced models. We release our code, model parameters, and new MIMIC-III and MIMIC-IV training and evaluation pipelines to accommodate fair future comparisons.\footnote{\label{footnote:source_code}\url{https://github.com/JoakimEdin/medical-coding-reproducibility}}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{paper_automated/mimiciv_code_freq.png}
    \caption[The frequency of ICD-9 and ICD-10 codes in MIMIC-IV before pre-processing.]{ The frequency of ICD-9 and ICD-10 codes in MIMIC-IV before pre-processing. As discussed in \cref{sec: splits}, we removed codes with fewer than ten instances (dashed line).}
    \label{fig:mimiciv_code_dist}
\end{figure}

@@

\textcite{mullenbachExplainablePredictionMedical2018} released code for pre-processing the discharge summaries, generating the train-test split, and evaluating model performance on MIMIC-III which many subsequent papers have used \parencite{vuLabelAttentionModel2020, huangPLMICDAutomaticICD2022, liICDCodingClinical2020, baoMedicalCodePrediction2021, yuanCodeSynonymsMatter2022, kimReadAttendCode2021}.
Pre-processing consisted of lower-casing all text and removing words that only contain out-of-alphabet characters. Predicting procedures and diagnosis codes were treated as a single task.
The dataset was split into training, validation, and test sets using random sampling, ensuring that no patient occurred in both the training and test set. The (non-stratified) random sampling lead to 54\% of the ICD codes in MIMIC-III \textit{full} not being sampled in the test set. This complicates the interpretation of results since these codes only contribute true negatives or false positives. 
Models are evaluated using the micro and macro average of the area under the curve of the receiver operating characteristics (AUC-ROC), F1-score, and precision@k.

While most papers use the pre-processing, train-test split, and evaluation metrics described above, they differ in several aspects of training. This may lead to performance differences unrelated to modeling choices which are undesirable when seeking to compare models. 
For instance, due to varying memory constraints of different models, documents are usually truncated to some maximum length. In the literature, this maximum varies from 2,500 to 4,000 words \parencite{mullenbachExplainablePredictionMedical2018, vuLabelAttentionModel2020, huangPLMICDAutomaticICD2022}.
Furthermore, not all papers tune the prediction decision boundary but simply set it to 0.5,
 hyperparameter search ranges and sampling methods  vary between works, and learning rate schedulers are only used in LAAT and PLM-ICD\parencite{mullenbachExplainablePredictionMedical2018, liICDCodingClinical2020}. In LAAT, the learning rate was decreased by 90\% when the F1-score had not increased for five epochs. PLM-ICD used a schedule with linear warm up followed by linear decay.

@@

\begin{sidewaystable}
    \centering
    \caption[Hyperparameters, maximum document lengths, and decision boundary tuning strategies used in the original works compared to improved settings.]{ Hyperparameters, maximum document lengths, and decision boundary tuning strategies used in the original works compared to the optimal settings found in this paper (marked with *). LR is the learning rate scheduler. ``Length" is the maximum number of words a document can contain before being truncated. $\dagger$ applies to models using word-piece tokenization. These models were filtered on the number of sub-words instead of full words. ``DB tune" is whether the optimal decision boundary was found using the validation set. If a paper did not tune the decision boundary, it was set to 0.5.}
    \label{tab:hyperparameters}
    \resizebox*{\textwidth}{!}{%
    \begin{tabular}{lccccccccc}
        \toprule
        & \multicolumn{7}{c}{Hyperparameters} & & \\
        \cmidrule(lr){2-8}
        Model & Batch Size & Weight Decay & Learning Rate & Dropout & LR Scheduler & Optimizer & Epochs & Length & DB tune \\
        \midrule
        Bi-GRU  & 16 &  0.0 & 0.003 &  0.2   & no  & Adam &  100 & 2500 & no\\
        Bi-GRU$^*$  & 8 &  0.0001 & 0.001 &  0   & yes  & AdamW &  20 & 4000 & yes\\
        \hline
        CNN  & 16 &  0.0 & 0.003  &  0.2 & no  & Adam &  100 & 2500 & no\\
        CNN$^*$    & 8 &  0.00001 & 0.001  &  0 & yes  & AdamW &  20 & 4000 & yes \\
        \hline
        CAML  & 16 &  0.0 & 0.0001  &  0.2  & no  & Adam  &  200 & 2500 & no\\
        CAML$^*$    & 8 &  0.001 & 0.005  &  0.2  & yes  & AdamW  &  20 & 4000 & yes\\
        \hline
        MultiResCNN & 16 &  0.0 & 0.0001  &  0.2  & no  & Adam &  200 & 2500 & no\\
        MultiResCNN$^*$  & 16 &  0.0001 & 0.0005  &  0.2  & yes  & AdamW &  20 & 4000 & yes\\
        \hline
        LAAT   & 8 &  0.0 & 0.0001  &  0.3  & yes  & AdamW &  50 & 4000 & no\\
        LAAT$^*$    & 8 &  0.001 & 0.001  &  0.2  & yes  & AdamW &  20 & 4000 & yes\\
        \hline
        PLM-ICD  & 8 &  0.0 & 0.00005  &  0.2  & yes  & AdamW &  20 & 3072$^\dagger$ & yes \\
        PLM-ICD$^*$   & 16 &  0.0 & 0.00005  &  0.2  & yes  & AdamW &  20 & 4000 & yes \\
        \bottomrule
    \end{tabular}%
    }
\end{sidewaystable}

Previous works calculated the macro F1-score as the harmonic mean of the macro precision and macro recall \parencite{mullenbachExplainablePredictionMedical2018, liICDCodingClinical2020, vuLabelAttentionModel2020, huangPLMICDAutomaticICD2022}. \textcite{opitzMacroF1Macro2021} analyze macro F1 formulas common in multi-class and multi-label classification. They demonstrate that the above formulation is suboptimal, as it rewards heavily biased classifiers in unbalanced datasets. Therefore, as recommended by the authors, we calculated the macro F1-score as the arithmetic mean of the F1-score for each class.
As seen in \cref{tab:subsets}, 54\% of codes in MIMIC-III \textit{full} are missing in the test set. Previous works set the F1-score of all the missing codes in the test set to 0, resulting in a misleadingly low macro F1-score. Because 54\% of the codes are missing, the maximum possible macro F1-score is 46\%. 
We ignored all codes not in the test set for our reproduction, essentially trading bias for variance. 
For our revised comparison, we resolved the issue by instead sampling new splits that reduce missing codes to a negligible fraction (see \cref{sec: splits}) and ignoring the few that were still missing.

@@

\subsection{Revised comparison}
To address the issues associated with comparing results reported by previous works described in \cref{sec: training and evaluation methods,sec:metrics}, we perform a revised model comparison. We run experiments on the new MIMIC-III \textit{clean}, MIMIC-IV \textit{ICD-9}, and \textit{ICD-10}.
All models were trained for 20 epochs using a learning rate schedule with linear warm up for the first 2K updates followed by linear decay \parencite{huangPLMICDAutomaticICD2022}. We found this schedule to speed up the training convergence of all the models.
Whereas original works use Adam or AdamW, we used AdamW for all experiments as it corrects the weight decay implementation of Adam \parencite{kingmaAdamMethodStochastic2017, loshchilovDecoupledWeightDecay2022}. For each model, we tuned the decision boundary to maximize the micro F1-score on the validation set. We used randomized sampling to find optimal settings for dropout, weight decay, learning rate, and batch size. The hyperparameter search was performed on MIMIC-III \textit{clean}, and the MIMIC-IV splits. We found that the best setting for each model generalized across datasets. Using this setting, we ran each model ten times with different seeds on each dataset. All documents were truncated to a maximum of 4000 words. The hyperparameters, maximum document lengths, and decision boundary tuning strategy are summarized in \cref{tab:hyperparameters}.

@@

\begin{sidewaystable}
    \centering
    \caption[Reproduced test set results compared with those from the original works.]{ Reproduced test set results compared with those from the original works. Our reproduced results are indicated with *. The results were reproduced on MIMIC-III v1.4 with the preprocessing pipeline and splits of \textcite{mullenbachExplainablePredictionMedical2018}. Each model was reproduced using the hyperparameters presented in the respective paper. We use both macro F1 formulas: Macro$^\dagger$ refers to the method used in the original work, while Macro refers to the corrected version used in this paper.}
    \label{tab:reproduced_results}
    \resizebox*{\textwidth}{!}{%
    \begin{tabular}{l  cc  ccc  cc  cc  ccc  c}
        \toprule
        & \multicolumn{7}{c}{MIMIC-III \textit{full}} & \multicolumn{6}{c}{MIMIC-III \textit{50}} \\
        \cmidrule(lr){2-8}\cmidrule(lr){9-14}
        & \multicolumn{2}{c}{AUC-ROC} & \multicolumn{3}{c}{F1} & \multicolumn{2}{c}{Precision@$k$}  & \multicolumn{2}{c}{AUC-ROC} & \multicolumn{3}{c}{F1} & Precision@$k$  \\
        & Micro & Macro & Micro & Macro$^\dagger$ & Macro & 8 & 15 & Micro & Macro & Micro & Macro$^\dagger$ & Macro & 5 \\
        \midrule
        CNN &  96.9 & 80.6 &  41.9& 4.2 & -& 58.1 &  48.8  & 90.7 & 87.6 & 62.5 & 57.6 & - & 62.0  \\
        CNN$^*$ & 97.3 & 83.1 & 41.5 & 3.4 & 6.7 & 61.9 & 47.2 & 91.9 & 89.2 & 64.9 & 58.8 & 58.0 &  62.6 \\
        \hline
        Bi-GRU & 97.1 & 82.2 & 41.7 & 3.8 & - & 58.5 & 44.5  &  89.2 & 82.8 & 54.9 &  48.4 & - &  59.1  \\
        Bi-GRU$^*$  & 98.0 & 87.1 & 42.6 & 3.6 & 7.0 & 65.0 & 49.8 & 89.3 & 85.2 & 56.1 & 46.2 & 43.1 & 57.9  \\
        \hline
        CAML  &  98.6 &  89.5 &  53.9 & 8.8 & - &  70.9 &  56.1  & 90.9 & 87.5 &  61.4 &  53.2 & - &   60.9  \\
        CAML$^*$ & 98.4 & 88.4 & 49.5 & 5.6 & 11.3 & 69.9 & 54.9 & 91.1 & 87.5 & 60.6 & 52.4 & 51.0 & 61.1  \\
        \hline
        MultiResCNN  &  98.6 &  91.0 & 55.2 & 8.6 & - & 73.4 & 58.4  &  93.8 &  89.9 &  67.0 &  60.6 & - &  64.1 \\
        MultiResCNN$^*$  & 98.6 & 90.8 & 56.5 & 9.2 & 18.5 & 73.4 & 58.4 & 92.4 & 89.7 & 67.3 & 62.2 & 61.1 & 63.4  \\
        \hline
        LAAT  &  98.8 &  91.9 &  57.5 & 9.9 & - &  74.5 &  59.1  &  94.6 &  92.5 &  71.5 &  66.6 & - &  67.5 \\
        LAAT$^*$ & 98.6 & 89.5 & 56.1 & 8.2 & 16.2 & 73.9 & 58.7 & 92.8 & 90.5 & 66.8 & 60.8 & 59.2 & 64.0  \\
        \hline
        PLM-ICD  &  98.9 & 92.6 &  59.8 & 10.4 & - &  77.1 &  61.3  & - & - & - & - & - & -  \\
        PLM-ICD$^*$ & 98.8 & 92.3 & 58.9 & 11.1 & 22.8 & 75.7 & 60.5 & 93.8 & 91.7 & 70.5 & 66.3 & 65.4 & 65.7 \\
        \bottomrule
    \end{tabular}%
    }
\end{sidewaystable}

@@

    \caption[Results on the MIMIC-III \textit{clean}, MIMIC-IV \textit{ICD-9} and MIMIC-IV  \textit{ICD-10} test sets presented as percentages.]{ Results on the MIMIC-III \textit{clean}, MIMIC-IV \textit{ICD-9} and MIMIC-IV  \textit{ICD-10} test sets presented as percentages. Micro F1-scores rank the table in ascending order. Each model was trained ten times with different seeds. We performed a McNemar's test with Bonferroni correction and found that all the models are significantly different ($p < 0.001$).}

@@

\begin{sidewaystable}
    \centering
    \caption[Ablation study on MIMIC-III \textit{clean}.]{ Ablation study on MIMIC-III \textit{clean}. The numbers are the micro/macro F1-scores on the test set.}
    \label{tab:ablation_study}
    \resizebox*{\textwidth}{!}{%
    \begin{tabular}{lcccccc}
        \toprule
         & PLM-ICD & LAAT & MultiResCNN & CAML & Bi-GRU & CNN \\
        \midrule
        Our result & 59.6/26.6 & 57.8/22.6 & 56.4/22.9 & 55.4/20.4 & 49.7/12.2 & 48.0/9.9 \\
        \hline
        Input length truncated at 2500 words & 59.4/26.2 & 57.6/22.3 & 56.0/23.2 & 54.8/19.7 & 49.4/12.0 & 47.9/9.8 \\
        No decision boundary tuning & 58.7/23.0 & 56.2/19.0 & 56.2/22.6 & 53.3/17.1 & 45.3/8.1 & 43.8/7.0 \\
        Original hyperparameters & 59.6/27.0 & 57.5/21.6 & 56.4/20.0 & 52.8/17.3 & 48.1/11.2 & 46.9/10.2 \\
        \bottomrule
        \end{tabular}%
    }
\end{sidewaystable}

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.9\textwidth}
        \centering
        \includegraphics[width=\textwidth]{paper_automated/threshold_tuning_micro_full.png}
        % \caption{}
        % \label{fig:threshold_tuning_micro}
    \end{subfigure}
    \begin{subfigure}[b]{0.9\textwidth}
        \centering
        \includegraphics[width=\textwidth]{paper_automated/threshold_tuning_macro_full.png}
        % \caption{}
        % \label{fig:threshold_tuning_macro}
    \end{subfigure}%
    \caption[Relationship between chosen threshold and F1-score of every reproduced model.]{ The relationship between chosen threshold and F1-score of every reproduced model in \cref{tab:reproduced_results}. The left figure shows the micro F1-score, and the right shows the macro F1-score. The models were evaluated on MIMIC-III \textit{clean}.}
    \label{fig:threshold_tuning}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.95\textwidth}
        \centering
        \includegraphics[width=\textwidth]{paper_automated/train_size_f1_micro.png}
        % \caption{}
        % \label{fig:train_size_micro}
    \end{subfigure}
    \begin{subfigure}[b]{0.95\textwidth}
        \centering
        \includegraphics[width=\textwidth]{paper_automated/train_size_f1_macro.png}
        % \caption{}
        % \label{fig:train_size_macro}
    \end{subfigure}
    \caption[The relationship between the number of training examples and F1-score on MIMIC-IV \textit{ICD-9}.]{  The relationship between the number of training examples and F1-score on MIMIC-IV \textit{ICD-9}. The left figure shows the F1 Micro score on the y-axis, while the right figure shows the F1 Macro score.}
    \label{fig:train_size}
\end{figure}

@@

\begin{table}
    \centering
    \caption[Correlation between the F1-score and the logarithm of code frequency and document length on MIMIC-IV \textit{ICD-9}.]{ Correlation between the F1-score and the logarithm of code frequency and document length on MIMIC-IV \textit{ICD-9}. As discussed in \cref{subsubsec:document-length}, we only considered document lengths between 1000 and 4000 words. All correlations are statistically significant ($p<0.001$).}
    \label{tab:correlations}
    \begin{tabular}{lcccc}
        \toprule
        & \multicolumn{2}{c}{Code frequency} & \multicolumn{2}{c}{Document lengths}\\
        \cmidrule(lr){2-3}\cmidrule(lr){4-5}
        & Pearson & Spearman & Pearson & Spearman \\
        \midrule
        CNN  & 0.61 & 0.68 & -0.09 & -0.08 \\
        Bi-GRU & 0.57 & 0.65 & -0.08 & -0.07 \\
        CAML & 0.56 & 0.60 & -0.03 & -0.03   \\
        MultiResCNN & 0.47 & 0.53 & -0.02 & -0.03  \\
        LAAT & 0.52 & 0.57 & -0.02 & -0.02 \\
        PLM-ICD & 0.48 & 0.52 & -0.02 & -0.02   \\
        \bottomrule
    \end{tabular}
\end{table}
